A Study in Pairwise Testing Heuristics for Granger-causal Network Estimation

### 2019-07-26 paper_appl/paper_appl.tex
Methods to compute causality graphs by different methods on the eeg data.

\section{Introduction and Review}
\label{sec:introduction}
Following recent theoretical developments \cite{my_GC_paper}
establishing that, under certain topological assumptions on the
underlying causality graph, it is possible to accurately recover the
true causality graph by way of pairwise causality tests alone.  That
is, there are algorithms which are immune to the effects of
confounders when applied to certain causality graph structures.  In
this paper we provide simulation evidence that, given structural
assumptions, such algorithms can dramatically outperform state of the
art LASSO-type estimators.  Moreover, even in cases where structural
assumptions do not in fact hold, using similar algorithms as pure
heuristics can still exhibit superior performance.  In addition to the
potential for gains in statistical power, that our algorithms make use
only of pairwise causality testing improves the scalability (both in
terms of the asymptotics, as well as through trivial parallelization)
of Granger-causality testing up to thousands of nodes.

@@

\subsection{Strongly Causal Graphs and Persistent Systems}
\label{sec:strongly_causal_graphs}
In this section we define the notion of a strongly causal graph, and
the implications for pairwise causality testing.

@@

  In order to exclude the possibility of certain types of cancellation
  in the $\VAR$ system of Equation \eqref{eq:var_model}, we require
  the system to have a property we have deemed ``persistence'' in
  \cite{my_GC_paper}, the intuition being that each node in the system
  maintains some ``memory''.

  \begin{definition}[Persistent]
    \hl{TODO:}I can provide the diagonalizability condition for
    $\VAR(p)$ models, but what about when $p = \infty$?
    \todo{persistence for $\VAR(\infty)$}
  \end{definition}

  The key property of persistent and strongly causal systems is that
  they are in some sense immune to confounders.  That is, if $i \pwgc j$
  \textit{and} $j \pwgc i$, then $i \ngc j$ and $j \ngc i$.  That is,
  bidirectional pairwise causation (which would otherwise be the
  result of hidden confounders) is not possible.

  \begin{theorem}[Pairwise Recovery \cite{my_GC_paper}]
    \label{thm:scg_recovery}
    If the Granger-causality graph $\gcg$ for persistent process
    $x(t)$ is a strongly causal DAG, then $\gcg$ can be inferred from
    pairwise causality tests, given an oracle for pairwise causation.

    \todo{Contain statement on convergence of the heuristic?}
    % \todo{Make this precisise} Moreover, given $T < \infty$ samples from
    % $x(t)$ then Algorithm \ref{alg:finite_pwgc} will correctly return
    % $\gcg$ with probability $1$ as $T \rightarrow \infty$.
  \end{theorem}

\section{Finite Sample Graph Recovery}
\label{sec:structure_learning}
In \cite{my_GC_paper} we prove Theorem \ref{thm:scg_recovery} by
providing an algorithm which, given an oracle for pairwise
Granger-causality, will correctly recover $\gcg$.  In this section,
we develop in detail appropriate statistical methods for
implementing this algorithm on a finite data sample.  When applied
in practice, the algorithm is a heuristic\footnote{Recall that
  Granger-causality is necessarily based on heuristics when $n$ is
  modestly large, as the complexity of searching over causality
  graphs scales as $O(2^{n^2})$}, since we do not know apriori
whether the causality graph of $x(t)$ is strongly causal (or if
$x(t)$ is persistent).  However, we will see in Section
\ref{sec:empirical_evaluation} that Algorithm \ref{alg:finite_pwgc}
outperforms LASSO-based heuristics, even when $\gcg$ does not
exactly satisfy the strong causality assumption.

@@

\section{Empirical Evaluation}
\label{sec:empirical_evaluation}
We have implemented our empirical experiments in Python \cite{scipy},
in particular we leverage the LASSO implementation from
\texttt{sklearn} \cite{sklearn} and the random graph generators from
\texttt{networkx} \cite{networkx}.  We run experiments using two
separate graph topologies having $n$ nodes: a strongly causal graph
(SCG) and a directed acyclic graph (DAG).  These are generated
respectively by drawing a random tree and a random Erdos Renyi graph
(with edge probability $q = \frac{2}{n}$ resulting in approximately
the same number of edges for the SCG as for the DAG), then creating a
directed graph by directing edges from lower numbered nodes to higher
numbered nodes.

We populate each of the edges (including self loops) with random
linear filters constructed by placing $5$ transfer function poles
(i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
guarantees stability for acyclic graphs).  The resulting system is
driven by i.i.d. Gaussian random noise, each component having random
variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
We set $p_{\text{max}} = 15$.  Results and representative graphs are
collected in Figures \ref{fig:random_graph_topologies},
\ref{fig:simulation_results_comparison1},
\ref{fig:simulation_results_comparison2},
\ref{fig:simulation_results_scaling_and_small_T},
\ref{fig:simulation_results_dense}.

We compare our results against the adaptive LASSO
\cite{adaptive_lasso_zou2006}, which outperformed substantially both
the LASSO and the grouped LASSO.  Motivated by scaling, we split the
squared error term into separate terms, one for each group of incident
edges on a node, and estimate the collection of $n$ incident filters
$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
$\xi_i^{\text{LASSO}}$ in the following:

\begin{equation}
  \begin{aligned}
  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
  \end{aligned}
\end{equation}

where we are choosing $\lambda$, the regularization parameters, via the BIC.

\begin{figure}
  \centering
  \caption{Representative Random Graph Topologies}
  \label{fig:random_graph_topologies}
  \begin{subfigure}[b]{0.3\textwidth}
    \caption{Random SCG}
    \includegraphics[width=\linewidth]{example_scg.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \caption{Random DAG $(q = \frac{2}{n})$}
    \includegraphics[width=\linewidth]{example_dag.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \caption{Random DAG $(q = \frac{4}{n})$}
    \includegraphics[width=\linewidth]{example_dag_q.pdf}
  \end{subfigure}
\end{figure}

\begin{remark}[Graph Topologies]
  We depict in Figure \ref{fig:random_graph_topologies} the topologies
  of random graphs used in our empirical evaluation.  For values of
  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
  a topology which is, at least qualitatively, close to the SCG.  As
  the value of $q$ increases, the random graphs deviate farther from
  the SCG topology, and we therefore expect the LASSO to outperform
  PWGC for larger values of $q$.  This can be observed in Figure
  \ref{fig:simulation_results_dense}, at least where performance is
  measured by the support recovery (i.e. via the MCC).
\end{remark}

\begin{remark}[MCC as a Support Recovery Measurement]
  We apply Matthew's Correlation Coefficient (MCC)
  \cite{matthews1975comparison} as a statistic for measuring support
  recovery performance.  This statistic synthesizes the confusion
  matrix into a single score appropriate for unbalanced labels and is
  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
  performance, $0$ being the performance of random guessing, and $-1$
  being perfectly opposed.
\end{remark}

\begin{remark}[Error Measurement]
  We estimate the 1-step ahead prediction error by forming the variance matrix estimate

  \begin{equation*}
    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
  \end{equation*}

  on a long stream of out-of-sample data.  We then report the quantity

  \begin{equation*}
    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
  \end{equation*}

  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
\end{remark}

\begin{figure}
  \centering
  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
  \label{fig:simulation_results_comparison1}
  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}

  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
    estimation.  We make comparisons against both the MCC and the
    relative log mean-squared prediction error
    $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
    in Figure \ref{fig:simulation_results_comparison1} are for systems
    guaranteed to satisfy the assumptions required for Theorem
    \ref{thm:scg_recovery}.}
\end{figure}

\begin{figure}
  \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
  \label{fig:simulation_results_comparison2}
  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}

  {\scriptsize Figure \ref{fig:simulation_results_comparison2}
    provides results for systems which do not guarantee the
    assumptions of Theorem \ref{thm:scg_recovery}, though the graph
    has a similar level of sparsity.}
\end{figure}

\begin{figure}
  \centering
  \caption{PWGC Scaling and Small Sample Performance}
  \label{fig:simulation_results_scaling_and_small_T}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Fixed $T$, increasing $n$ (SCG)}
    \label{fig:simulation_results_scaling}
    \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{MCC Comparison for $T \le 100$}
    \label{fig:small_T_comparison}
    \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
  \end{subfigure}

  {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
    support recovery performance as the number of nodes $n$ increases,
    and the edge proportion as well as the number of samples $T$ is
    held fixed.  Remarkably, the degradation as $n$ increases is
    limited, it is primarily the graph topology (SCG or non-SCG) as
    well as the level of sparsity (measured by $q$) which are the
    determining factors for support recovery performance.

    Figure \ref{fig:small_T_comparison} provides a support recovery
    comparison for very small values of, $T$ typical for many
    applications.}
\end{figure}

\begin{figure}
  \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
  \label{fig:simulation_results_dense}
  \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}

  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
    comparison between PWGC and AdaLASSO as the density of graph edges
    (as measured by $q$) increases.  For reference,
    $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
    as the SCGs we simulated.  As $q$ increases, the AdaLASSO
    outperforms PWGC as measured by the MCC.  However, PWGC maintains
    superior performance for 1-step-ahead prediction.  We speculate
    that this is a result of fitting the sparsity pattern recovered by
    PWGC via OLS which directly seeks to optimize this metric, whereas
    the LASSO is encumbered by the sparsity inducing penalty.}
\end{figure}

In reference to figure \ref{fig:simulation_results_comparison1} it
should not be overly surprising that our PWGC algorithm performs
better than the LASSO for the case of a strongly causal graph, since
in this case the assumptions which guarantee the correctness of
Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
still markedly superior in the case of a more general DAG.  We would
conjecture that a DAG having a similar degree of sparsity as an SCG is
likely to be ``close'' to an SCG.  Figure
\ref{fig:simulation_results_dense} illustrates the severe (expected)
degradation in performance as the number of edges increases while
the number of data samples $T$ remains fixed.  For larger values $q$
in this plot, the number of edges in the graph is comparable to the
number of data samples.

We have also paid close attention to the performance of PWGC in the
very small sample ($T \le 100$) regime (see Figure
\ref{fig:small_T_comparison}), as this is the regime many applications
must contend with.

In regards scalability, we have observed that performing the $O(n^2)$
pairwise Granger-causality calculations consumes the vast majority
($> 90\%$) of the computation time.  Since this step is trivially
parallelizable, our algorithm also scales well with multiple cores or
multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
demonstration of this scalability, where we are able to estimate
graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
using only $T = 500$ data points, granted, an SCG on this many nodes
is extremely sparse.

\begin{table}
  \centering
  \caption{Simulation Results: PWGC vs AdaLASSO}
  \label{taab:simulation_table}

  \begin{tabular}{|ll||ll|ll|ll|}
    \toprule
    &{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
    &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
    \textbf{T}&\textbf{q}&&&&&&\\
    \midrule
    \multirow{4}{*}{\textbf{50}}
    &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
    &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
    &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
    &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
    \midrule
    \multirow{4}{*}{\textbf{250}}
    &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
    &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
    &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
    &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
    \midrule
    \multirow{4}{*}{\textbf{1250}}
    &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
    &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
    &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
    &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
    \bottomrule
  \end{tabular}

  {\scriptsize Results of Monte Carlo simulations comparing PWGC and
    AdaLASSO for small samples and when the SCG assumption doesn't
    hold.  The superior result is bolded when the difference is
    statistically significant, as measured by
    \texttt{scipy.stats.ttest\_rel}.  100 iterations are run for each
    set of parameters.}
\end{table}

\section{Application}
\label{sec:application}
In this section we apply our methods to a real set of EEG data
obtained from the ``EEG Database Data
Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
\cite{zhang1995event} on the UCI machine learning repository
\cite{uci_mlr}.  This dataset contains 1 second long measurements of
(64 channel) EEG signals from patients who are given visual stimuli.
The subjects in the study are labeled as being either ``control'' or
``alcoholic'', however, we will ignore this label and instead focus on
\textit{distinguishing between subjects} based on the
Granger-causality graphs inferred from the subject's trials.  Our
reasoning is to focus on the underlying question: ``Does our PWGC
algorithm uncover meaningful Granger-causal connections from EEG
data?''.  Focusing only on the subject label allows us to answer this
question without simultaneously grappeling with the physiological
question of whether alcoholic subjects as a group have discernable
differences in their EEG readings\footnote{Some exploratory analysis
  actually suggests that alcoholic Granger-causality graphs are not
  substantively different}, which is a stronger requirement than
simply that there are meaningful distinctions between the EEG readings
of subjects generally.

The dataset consists of
$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
$N = 10723$.  There are on average 90 trials for each subject,
ranging between 30 and 119.

We have constructed a simple pipeline for discriminating between
subjects by first applying an iterative PWGC algorithm (see Algorithm
\ref{alg:iterated_pwgc}) directly to the EEG data $x^{(i)}$ to obtain
a Granger-causality graph $\widehat{G}^{(i)}$.  We then feed the
vectorized adjacency matrix of $\widehat{G}^{(i)}$ through a
polynomial Kernel multinomial logistic regression model with
parameters fit by cross validation.

\begin{remark}
  The iterated application of PWGC is an example of a heuristic by
  which graph estimates that aren't constrained to being strongly
  causal can be obtained.  We defer to future work the theoretical
  analysis of more sophisticated heuristics, or algorithms appropriate
  for different topological assumptions.
\end{remark}

@@

While it is almost certainly possible to achieve much greater
classification accuracy on this dataset by constructing a classifier
to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
relevant information is being captured by the graph estimates
$\widehat{G}^{(i)}$.  That is, that the latent structure
$\widehat{G}^{(i)}$ may provide scientifically relevant insights, as
opposed to simply being a feature for classification tasks.

Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_error}
illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
as opposed simply to a collection of $n$ unidimensional
autoregressiive models.  The conclusion being that the
Granger-causality graph estimates are not purely spurious.

>>>>

\section{Introduction and Review}
\label{sec:introduction}
Following recent theoretical developments \cite{my_GC_paper}
establishing that, under certain topological assumptions on the
underlying causality graph, it is possible to accurately recover the
true causality graph by way of pairwise causality tests alone.  That
is, there are algorithms which are immune to the effects of
confounders when applied to certain causality graph structures.  See
\cite{tam2013gene_pwgc} for earlier work on pairwise Granger-causality.
In this paper we provide simulation evidence that, given structural
assumptions, such algorithms can dramatically outperform state of the
art LASSO-type estimators.  Moreover, even in cases where structural
assumptions do not in fact hold, using similar algorithms as pure
heuristics can still exhibit superior performance.  In addition to the
potential for gains in statistical power, that our algorithms make use
only of pairwise causality testing improves the scalability (both in
terms of the asymptotics, as well as through trivial parallelization)
of Granger-causality testing up to thousands of nodes.

@@

\subsection{Exact Recovery From Pairwise Tests}
\label{sec:strongly_causal_graphs}
In this section we define the notion of a strongly causal graph, and
review the implications for pairwise causality testing.

@@

  There is another technical condition (detailed in
  \cite{my_GC_paper}) required for the next theorem which eliminates
  the possibility of certain kinds of pathological cancellation when
  confounding nodes are present.  However, $\VAR$ systems which do not
  satisfy the technical condition are pathological.

  \begin{theorem}[Pairwise Recovery \cite{my_GC_paper}]
    \label{thm:scg_recovery}
    Suppose that the Granger-causality graph $\gcg$ for the process
    $x(t)$ is a strongly causal DAG.  Then, except for certain
    pathological cases \cite{my_GC_paper}, $\gcg$ can be inferred from
    pairwise causality tests, given an oracle for pairwise causation.
  \end{theorem}

  The proof of this theorem proceeds by proving the correctness of an
  algorithm for implementing exact recovery through pairwise testing.
  This algorithm inspires the heuristic\footnote{Recall that
    Granger-causality is necessarily based on heuristics when $n$ is
    modestly large, as the complexity of searching over causality
    graphs scales as $O(2^{n^2})$} of Section
  \cite{sec:structure_learning}.

\section{Finite Sample Graph Recovery}
\label{sec:structure_learning}
In this section we detail our proposed pairwise testing heuristics.
In particular, Algorithm \ref{alg:finite_pwgc} proceeds first by a
simple thresholding of pairwise relations.  The algorithm next
proceeds by an iterative procedure which ``peels away layers'' of the
causality graph eminating from parentless nodes, while maintaining the
strong causality property.  We present simulation evidence that this
heuristic can outperform LASSO-based heuristics in \cite{my_GC_paper}.
% We will see in Section \ref{sec:empirical_evaluation} that Algorithm
% \ref{alg:finite_pwgc} outperforms LASSO-based heuristics, even when
% $\gcg$ does not exactly satisfy the strong causality assumption.

@@

In order to adapt Algorithm \ref{alg:finite_pwgc} to graphs which may
not be strongly causal, we consider an iterative application as
described in Algorithm \ref{alg:iterated_pwgc}.  We apply Algorithm
\ref{alg:finite_pwgc} first to the data itself, and then subsequently
to the resulting residuals before performing a final estimate on the
combined graph.  We evaluate these algorithms, as well as simpler
pairwise heuristics and LASSO-type heuristics in the next section.  A
theoretical analysis of Algorithm \ref{alg:iterated_pwgc} is deferred
to future work.

@@

% \section{Empirical Evaluation}
% \label{sec:empirical_evaluation}
Our implementations are in Python \cite{scipy}, in particular we
leverage the LASSO implementation from \texttt{sklearn} \cite{sklearn}
and the random graph generators from \texttt{networkx}
\cite{networkx}.  Our paper \cite{my_GC_paper} provides simulation
evidence that our PWGC heuristic can outperform LASSO-based heuristics
for simulated data.

% We run experiments using two separate graph topologies having $n$
% nodes: a strongly causal graph (SCG) and a directed acyclic graph
% (DAG).  These are generated respectively by drawing a random tree
% and a random Erdos Renyi graph (with edge probability
% $q = \frac{2}{n}$ resulting in approximately the same number of
% edges for the SCG as for the DAG), then creating a directed graph by
% directing edges from lower numbered nodes to higher numbered nodes.

% We populate each of the edges (including self loops) with random
% linear filters constructed by placing $5$ transfer function poles
% (i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
% guarantees stability for acyclic graphs).  The resulting system is
% driven by i.i.d. Gaussian random noise, each component having random
% variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
% We set $p_{\text{max}} = 15$.  Results and representative graphs are
% collected in Figures \ref{fig:random_graph_topologies},
% \ref{fig:simulation_results_comparison1},
% \ref{fig:simulation_results_comparison2},
% \ref{fig:simulation_results_scaling_and_small_T},
% \ref{fig:simulation_results_dense}.

% We compare our results against the adaptive LASSO
% \cite{adaptive_lasso_zou2006}, which outperformed substantially both
% the LASSO and the grouped LASSO.  Motivated by scaling, we split the
% squared error term into separate terms, one for each group of incident
% edges on a node, and estimate the collection of $n$ incident filters
% $\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
% $\xi_i^{\text{LASSO}}$ in the following:

% \begin{equation}
%   \begin{aligned}
%   \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
%   \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
%   \end{aligned}
% \end{equation}

% where we are choosing $\lambda$, the regularization parameters, via the BIC.

% \begin{figure}
%   \centering
%   \caption{Representative Random Graph Topologies}
%   \label{fig:random_graph_topologies}
%   \begin{subfigure}[b]{0.3\textwidth}
%     \caption{Random SCG}
%     \includegraphics[width=\linewidth]{example_scg.pdf}
%   \end{subfigure}
%   \begin{subfigure}[b]{0.3\textwidth}
%     \caption{Random DAG $(q = \frac{2}{n})$}
%     \includegraphics[width=\linewidth]{example_dag.pdf}
%   \end{subfigure}
%   \begin{subfigure}[b]{0.3\textwidth}
%     \caption{Random DAG $(q = \frac{4}{n})$}
%     \includegraphics[width=\linewidth]{example_dag_q.pdf}
%   \end{subfigure}
% \end{figure}

% \begin{remark}[Graph Topologies]
%   We depict in Figure \ref{fig:random_graph_topologies} the topologies
%   of random graphs used in our empirical evaluation.  For values of
%   $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
%   a topology which is, at least qualitatively, close to the SCG.  As
%   the value of $q$ increases, the random graphs deviate farther from
%   the SCG topology, and we therefore expect the LASSO to outperform
%   PWGC for larger values of $q$.  This can be observed in Figure
%   \ref{fig:simulation_results_dense}, at least where performance is
%   measured by the support recovery (i.e. via the MCC).
% \end{remark}

% \begin{remark}[MCC as a Support Recovery Measurement]
%   We apply Matthew's Correlation Coefficient (MCC)
%   \cite{matthews1975comparison} as a statistic for measuring support
%   recovery performance.  This statistic synthesizes the confusion
%   matrix into a single score appropriate for unbalanced labels and is
%   calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
%   performance, $0$ being the performance of random guessing, and $-1$
%   being perfectly opposed.
% \end{remark}

% \begin{remark}[Error Measurement]
%   We estimate the 1-step ahead prediction error by forming the variance matrix estimate

%   \begin{equation*}
%     \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
%   \end{equation*}

%   on a long stream of out-of-sample data.  We then report the quantity

%   \begin{equation*}
%     \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
%   \end{equation*}

%   where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
% \end{remark}

% \begin{figure}
%   \centering
%   \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
%   \label{fig:simulation_results_comparison1}
%   \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}

%   {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
%     estimation.  We make comparisons against both the MCC and the
%     relative log mean-squared prediction error
%     $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
%     in Figure \ref{fig:simulation_results_comparison1} are for systems
%     guaranteed to satisfy the assumptions required for Theorem
%     \ref{thm:scg_recovery}.}
% \end{figure}

% \begin{figure}
%   \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
%   \label{fig:simulation_results_comparison2}
%   \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}

%   {\scriptsize Figure \ref{fig:simulation_results_comparison2}
%     provides results for systems which do not guarantee the
%     assumptions of Theorem \ref{thm:scg_recovery}, though the graph
%     has a similar level of sparsity.}
% \end{figure}

% \begin{figure}
%   \centering
%   \caption{PWGC Scaling and Small Sample Performance}
%   \label{fig:simulation_results_scaling_and_small_T}
%   \begin{subfigure}[b]{0.45\textwidth}
%     \caption{Fixed $T$, increasing $n$ (SCG)}
%     \label{fig:simulation_results_scaling}
%     \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
%   \end{subfigure}
%   \begin{subfigure}[b]{0.45\textwidth}
%     \caption{MCC Comparison for $T \le 100$}
%     \label{fig:small_T_comparison}
%     \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
%   \end{subfigure}

%   {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
%     support recovery performance as the number of nodes $n$ increases,
%     and the edge proportion as well as the number of samples $T$ is
%     held fixed.  Remarkably, the degradation as $n$ increases is
%     limited, it is primarily the graph topology (SCG or non-SCG) as
%     well as the level of sparsity (measured by $q$) which are the
%     determining factors for support recovery performance.

%     Figure \ref{fig:small_T_comparison} provides a support recovery
%     comparison for very small values of, $T$ typical for many
%     applications.}
% \end{figure}

% \begin{figure}
%   \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
%   \label{fig:simulation_results_dense}
%   \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}

%   {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
%     comparison between PWGC and AdaLASSO as the density of graph edges
%     (as measured by $q$) increases.  For reference,
%     $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
%     as the SCGs we simulated.  As $q$ increases, the AdaLASSO
%     outperforms PWGC as measured by the MCC.  However, PWGC maintains
%     superior performance for 1-step-ahead prediction.  We speculate
%     that this is a result of fitting the sparsity pattern recovered by
%     PWGC via OLS which directly seeks to optimize this metric, whereas
%     the LASSO is encumbered by the sparsity inducing penalty.}
% \end{figure}

% In reference to figure \ref{fig:simulation_results_comparison1} it
% should not be overly surprising that our PWGC algorithm performs
% better than the LASSO for the case of a strongly causal graph, since
% in this case the assumptions which guarantee the correctness of
% Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
% still markedly superior in the case of a more general DAG.  We would
% conjecture that a DAG having a similar degree of sparsity as an SCG is
% likely to be ``close'' to an SCG.  Figure
% \ref{fig:simulation_results_dense} illustrates the severe (expected)
% degradation in performance as the number of edges increases while
% the number of data samples $T$ remains fixed.  For larger values $q$
% in this plot, the number of edges in the graph is comparable to the
% number of data samples.

% We have also paid close attention to the performance of PWGC in the
% very small sample ($T \le 100$) regime (see Figure
% \ref{fig:small_T_comparison}), as this is the regime many applications
% must contend with.

% In regards scalability, we have observed that performing the $O(n^2)$
% pairwise Granger-causality calculations consumes the vast majority
% ($> 90\%$) of the computation time.  Since this step is trivially
% parallelizable, our algorithm also scales well with multiple cores or
% multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
% demonstration of this scalability, where we are able to estimate
% graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
% using only $T = 500$ data points, granted, an SCG on this many nodes
% is extremely sparse.

% \begin{table}
%   \centering
%   \caption{Simulation Results: PWGC vs AdaLASSO}
%   \label{taab:simulation_table}

%   \begin{tabular}{|ll||ll|ll|ll|}
%     \toprule
%     &{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
%     &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
%     \textbf{T}&\textbf{q}&&&&&&\\
%     \midrule
%     \multirow{4}{*}{\textbf{50}}
%     &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
%     &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
%     &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
%     &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
%     \midrule
%     \multirow{4}{*}{\textbf{250}}
%     &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
%     &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
%     &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
%     &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
%     \midrule
%     \multirow{4}{*}{\textbf{1250}}
%     &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
%     &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
%     &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
%     &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
%     \bottomrule
%   \end{tabular}

%   {\scriptsize Results of Monte Carlo simulations comparing PWGC and
%     AdaLASSO for small samples and when the SCG assumption doesn't
%     hold.  The superior result is bolded when the difference is
%     statistically significant, as measured by
%     \texttt{scipy.stats.ttest\_rel}.  100 iterations are run for each
%     set of parameters.}
% \end{table}

\section{Application}
\label{sec:application}
In this section we apply our methods, as well as a number of
previously known methods, to a real set of EEG data obtained from the
``EEG Database Data
Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
\cite{zhang1995event} on the UCI machine learning repository
\cite{uci_mlr}.  This dataset contains 1 second long measurements of
(64 channel) EEG signals from patients who are given visual stimuli.

More precisely, our dataset is described by
$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
$N = 10723$.  There are on average 90 trials for each subject, ranging
between 30 and 119.
